DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report

### 2022-10-10 Draft-Ped-Metrics-Example.Rmd
handle shifted children and clean up score tables

library(kableExtra)

@@

outline <- sf::read_sf(paste0(this_dir,"data/outline"))
outline$plain = 1
kitsap <- ggplot() + 
  geom_sf(data = outline["plain"]) + 
  theme_void()

@@

@@

colnames(groupcl) <- c("Road Classification", 
                       "Count of Roads", "Count of Segments", 
                       "Length in Miles") 
knitr::kable(groupcl, 
             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
  kable_material(c("striped", "hover"))
```

@@

knitr::kable(df, 
             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
  kable_material(c("striped", "hover"))
```

@@

knitr::kable(df, 
             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
  kable_material(c("striped", "hover"))
```

```{r}
# tmp_sw_sh %>%
#   st_set_geometry(., NULL) %>%
#   group_by(RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, TYPE) %>% # road_type) %>%
#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280) %>%
#   ungroup() %>%
#   pivot_wider(names_from = SIDE_OF_RO, values_from = "TYPE")
#%>%
  #mutate(both = !is.na(RIGHT) & !is.na(LEFT)) %>%
  #group_by(both) %>%
  #summarise(n = n())

# tmp_sw_ncdf %>%  
#   st_set_geometry(., NULL) %>%
#   filter(LENGTH > 0) %>%
#   group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
# 
# tmp_sh_ncdf %>%  
#   st_set_geometry(., NULL) %>%
#   filter(LENGTH > 0) %>%
#   group_by(SIDE_OF_RO, WIDTH) %>% # road_type) %>%
#   summarise(miles = sum(LENGTH, na.rm=TRUE)/5280)
```

@@

```{r combine-rd-sw-sh}
# combine the sidewalk and shoulder and roads data
x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
x$RD_LOG_ID <- 43809
comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
                                     SEGMENT_ID_R = SEGMENT_ID,
                                     FULL_NAME_R = FULL_NAME, 
                                     LENGTH_R = LENGTH,
                                     road_class, RUCODE), 
                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
  mutate(distance = as.integer(st_length(.)))

@@

knitr::kable(example_tbl) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "float_left")
```

@@

knitr::kable(score_interpretation,
             caption = "Interpretation of Scores for Segments and for Areas")  %>%
   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

```{r}
comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
  left_join(comb_score) %>%
  mutate(the_score = ifelse(is.na(score), 0, score),
         the_length = ifelse(is.na(distance), LENGTH, distance),
         w_score = the_score*the_length)

@@

```{r}
knitr::kable(comb_score_by_setting_class) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

Some of the segments with score of 0 are because they have shoulders when they should have sidewalks. If you are familiar with Central Kitsap, you might recognize this on Provost - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate. Here's a few example segments.

@@

Some road segments are missing entirely from the centerline road data - for example, the portion of Rigetop that goes from the end of Myhre and curves back to Silverdale Way. These missing road segments are mostly complete streets with wide sidewalks. Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre. These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.

@@

We can also look at regions within the County. For these, we identify the roads and any sidewalks or shoulders.^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]

@@

These regions and pedestrian generators are:
 * Commissioner Districts
 * UGAs
 * LAMIRDs
 * Schools
 * Libraries
 * Transit Centers

```{r}
score_breakdowns <- tibble(Area = "Kitsap County", Score = comb_total_score)
```

### Commissioner Districts

```{r plot_commdists}
commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
commdists <- st_as_sf(commdists) %>% 
  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
                ifelse(DISTRICT == "2", "2 South", "3 Central")))

kitsap +
  geom_sf(data = commdists, aes(fill = NAME), alpha = 0.5) +
  theme_void() +
  labs(title = 'Kitsap County Commissioner Districts',
       caption = "Data source: Kitsap GIS - data in development",
       color = "DIMS")

```{r loop_commdists,include=FALSE,echo=FALSE}
the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
loop_df <- st_as_sf(commdists) %>% mutate(tomatch = NAME)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_commdists, Score = NA))

expanded_child <- lapply(
    the_commdists,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )

    }
    )

parsed_child <- knitr::knit_child(text = unlist(expanded_child))

Kitsap County has `r length(the_commdists)` Commissioner Districts.

`r parsed_child`

### UGAs

```{r plot_ugas}
uga_no_city <- st_as_sf(uga) %>%
  filter(substring(GMA_JURISD, 1, 4) != "City")
kitsap +
  geom_sf(data = uga_no_city, aes(fill = GMA_JURISD), alpha = 0.5) +
  theme_void() +
  labs(title = 'Kitsap County UGAs',
       caption = "Data source: Kitsap GIS - data in development",
       color = "UGAS")

```{r loop_ugas,include=FALSE,echo=FALSE}
the_ugas <- unique(st_set_geometry(uga_no_city, NULL)$GMA_JURISD)
loop_df <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_ugas, Score = NA))

expanded_child <- lapply(
    the_ugas,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )

    }
    )

parsed_child <- knitr::knit_child(text = unlist(expanded_child))

Kitsap County has `r length(the_ugas)` UGAs - Urban Growth Areas that are outside of cities. Within the Urban Growth Areas, most design and planning standards are similar to what would be expected in an urban environment for new development.

`r parsed_child`

### LAMIRDs

```{r plot_lamirds}
kitsap +
  #geom_sf(data = nc_sw_sh, aes(color = as.factor(TYPE))) + 
  geom_sf(data = lamird, aes(fill = PLAN_AREA), alpha = 0.5) +
  theme_void() +
  labs(title = 'Kitsap County LAMIRDS',
       caption = "Data source: Kitsap GIS - data in development",
       color = "LAMIRDS")

```{r loop_lamirds,include=FALSE,echo=FALSE}
the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
loop_df <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_lamirds, Score = NA))

expanded_child <- lapply(
    the_lamirds,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )

    }
    )

parsed_child <- knitr::knit_child(text = unlist(expanded_child))

```
Kitsap County has `r length(the_lamirds)` LAMIRDS - Local Area of More Intense Rural Development.

`r parsed_child`

```{r}
getWalkingPolygonInCounty <- function(locdf, idcol) {
  # finds the items in locdf that are in the target area
  # then finds the polygon around them
  # then limits that again to the area in the target (because the polygon may go into a city)
  if (!is.na(st_crs(locdf)$epsg)) { # the locdf has a epsg and cnc is NA
    cnc <- st_transform(county_no_cities, st_crs(locdf)$epsg)
    fdf <- st_intersection(locdf, st_buffer(cnc,0)) 
  } else {
    fdf <- st_intersection(locdf, st_buffer(county_no_cities,0)) 
  }
  fdf2 <- mapboxapi::mb_isochrone(fdf,
                                  time = 30, "walking", id_column=idcol)
  cnc <- st_transform(county_no_cities, st_crs(fdf2)$epsg)
  st_intersection(fdf2, st_buffer(cnc,0)) 
}
```

### Schools
```{r}
pub_schools <- schools %>%
  filter(TYPE == "PUBLIC")

pub_schools2 <- getWalkingPolygonInCounty(pub_schools, "NAME")
```

Kitsap County has `r nrow(pub_schools)` Public schools; however, only `r nrow(pub_schools2)` are outside of city limits. Here they are shown with a 30 minute walking distance.

```{r plot_pubschools}
#| fig.width = 8,
#| fig.height = 8,
#| out.width = '100%',
#| dpi = 96

kitsap +
  geom_sf(data = pub_schools2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = pub_schools, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County Public Schools',
       subtitle = "Schools not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap GIS - data in development\n
                  Walking distance polygons from Mapbox",
       fill = "Public Schools")

```{r loop_pubschools,include=FALSE,echo=FALSE, cache=TRUE}
the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
loop_df <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_schools, Score = NA))

expanded_child <- lapply(
    the_schools,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )
    })

parsed_child <- knitr::knit_child(text = unlist(expanded_child))
```

`r parsed_child`

### Libraries

```{r}
libraries <- community %>%
  filter(TYPE == "PUBLIC LIBRARY")

lib2 <- getWalkingPolygonInCounty(libraries, "NAME")
```

Kitsap County has `r nrow(libraries)` Public libraries; however, only `r nrow(lib2)` are outside of city limits. Here they are shown with a 30 minute walking distance.

```{r plot_libraries}
kitsap +
  geom_sf(data = lib2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = libraries, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County libraries',
       subtitle = "Libraries not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap GIS - data in development\n
                  Walking distance polygons from Mapbox",
       fill = "Libraries")

```{r loop_libraries,include=FALSE,echo=FALSE, cache = TRUE}
the_libraries <- unique(st_set_geometry(lib2, NULL)$id)
loop_df <- st_as_sf(lib2) %>% mutate(tomatch = id)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_libraries, Score = NA))

expanded_child <- lapply(
    the_libraries,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )
    })

parsed_child <- knitr::knit_child(text = unlist(expanded_child))
```

`r parsed_child`

### Transit

According to Kitsap Transit, "Transit Centers and Transfer Points are locations where Kitsap Transit buses and ferries meet up to allow the safe transfer of passengers from one route to another. "[@ktransit]^[Transit Center location data were manually copied from the Kitsap Transit Website because they are not one of the identified data sets on the Kitsap County GIS site. Future users will need to ensure that these addresses are correct.]

```{r}
transit_cp <- tribble(
  ~NAME, ~ADDRESS,
  "Bremerton Transportation Center (Ferry Terminal)","10 Washington Ave, Bremerton 98337",
  "Bainbridge Island Ferry Terminal","270 Olympic Drive SE, Bainbridge 98110",
  "Wheaton Way Transit Center","3915 Wheaton Way, Bremerton 98310",
  "West Bremerton Transit Center","540 Bruenn Ave, Bremerton 98312",
  "Silverdale Transit Center","Greaves Way & Kitsap Mall Blvd, Silverdale 98383",
  "North Viking Transit Center","21992 Viking Ave NW, Poulsbo 98370",
  "Kingston Ferry Terminal","11264 State Route 104, Kingston 98346",
  "Port Orchard Ferry Dock","73 Sidney Ave, Port Orchard 98366",
  "Southworth Ferry Terminal","11564 SE State Hwy. 160, Southworth 98386",
  "Annapolis Ferry Dock","1076 Bay Street, Port Orchard 98366",
  "Hwy. 305 & Suquamish Way","16003 WA-305, Poulsbo, Washington 98370",
  "McWilliams Park & Ride","1601 NE McWilliams Rd, Bremerton, Washington 98311",
  "Suquamish Park & Ride","18829 Division Ave NE, Suquamish, Washington 98392",
  "Port Orchard Wal-Mart","3497 Bethel Rd SE, Port Orchard 98366",
  "George's Corner Park & Ride","27618 Hansville Rd NE, Kingston 98346",
  "Gateway Fellowship Park & Ride","18901 8th Ave NE, Poulsbo 98370",
  "Miller Bay & Indianola","23404 Miller Bay Rd NE Poulsbo, Washington 98370",
  "St. Gabriel's Church","1150 Mitchell Ave SE, Port Orchard 98366"
)
# geocode the addresses into geocode = c(lon, lat)
transit_cp <- transit_cp %>% #tmaptools::geocode_OSM(transit_cp$ADDRESS)
  rowwise() %>%
  mutate(geocode = list(set_names(mapboxapi::mb_geocode(ADDRESS),
                        c("lon","lat")))) %>%
  unnest_wider(geocode)

# convert the lat lon into a point geometry
transit_cp <- st_as_sf(transit_cp, 
                      coords = c(x = "lon", y = "lat"), 
                      crs = 4326) 
transit_cp2 <- getWalkingPolygonInCounty(transit_cp, "NAME")
```

Kitsap County has `r nrow(transit_cp)` Public libraries; however, only `r nrow(transit_cp2)` are outside of city limits. Here they are shown with a 30 minute walking distance.

```{r plot_transit}
kitsap +
  geom_sf(data = transit_cp2, aes(fill = id), 
          lwd = 0, alpha = 0.3) +
  geom_sf(data = transit_cp, aes(color = 1), color = "black", show.legend = FALSE) +
  theme_void() +
  labs(title = 'Kitsap County Transit Centers and Transfer Points',
       subtitle = "Centers not labeled with color are not in the unincorporated county",
       caption = "Data source: Kitsap Transit Website\n
                  Walking distance polygons from Mapbox",
       fill = "Transit Centers")

```{r loop_transit,include=FALSE,echo=FALSE, cache = TRUE}
the_transits <- unique(st_set_geometry(transit_cp2, NULL)$id)
loop_df <- st_as_sf(transit_cp2) %>% mutate(tomatch = id)

score_breakdowns <- score_breakdowns %>%
  bind_rows(tibble(Area = the_transits, Score = NA))

expanded_child <- lapply(
    the_transits,
    function(x) {
      knitr::knit_expand(file = "child_single_geo.Rmd",
                                   current_geo = x )
    })

parsed_child <- knitr::knit_child(text = unlist(expanded_child))
```

`r parsed_child`

## Conclusion

@@

Smaller breakdowns of the county fare differently.

@@

getScoreBreakdownForArea <- function(the_area) {
  knitr::kable(score_breakdowns %>% 
                 filter(Area %in% eval(as.symbol(the_area))) %>%
                 mutate(Score = nice_comma(Score,2) 
                                           ),
             caption = paste("Scores for",
                             str_to_title(sub("the_"," ",the_area))
                             )
             ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
}

@@

```{r, echo=FALSE, results='asis'}
score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
res <- lapply(c("the_commdists","the_ugas", "the_lamirds", "the_schools", "the_libraries", "the_transits"), function(x) {
  knitr::knit_child(text = c(
    '## `r str_to_title(gsub("_"," ",x))`',
    '',
    '```{r}',
    'getScoreBreakdownForArea(x)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')

@@

 * missing part of Ridgetop Blvd in Silverdale where sidewalks are known to exist
 * unmatched road log identifiers between road and sidewalk and shoulder data

@@

In addition, many road segments have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?**

@@

## References

>>>>

library(kableExtra)
library(tidyverse)
library(sf)
```

@@

outline <- sf::read_sf(paste0(this_dir,"data/outline"))
outline$plain = 1
kitsap <- ggplot() + 
  geom_sf(data = outline["plain"], alpha = 0.5) + 
  theme_void()

@@

```{r gwfunction}
getWalkingPolygonInCounty <- function(locdf, idcol) {
  # finds the items in locdf that are in the target area
  # then finds the polygon around them
  # then limits that again to the area in the target (because the polygon may go into a city)
  if (!is.na(st_crs(locdf)$epsg)) { # the locdf has a epsg and cnc is NA
    cnc <- st_transform(county_no_cities, st_crs(locdf)$epsg)
    fdf <- st_intersection(locdf, st_buffer(cnc,0)) 
  } else {
    fdf <- st_intersection(locdf, st_buffer(county_no_cities,0)) 
  }
  fdf2 <- mapboxapi::mb_isochrone(fdf,
                                  time = 30, "walking", id_column=idcol)
  cnc <- st_transform(county_no_cities, st_crs(fdf2)$epsg)
  st_intersection(fdf2, st_buffer(cnc,0)) 
}
```

```{r smallerdata}
# this section sets up the data that will be used for subdivisions of the county
# get any data not set up yet
commdists <- sf::read_sf(paste0(this_dir,"data/commdist"))
commdists <- st_intersection(commdists, st_buffer(county_no_cities,0)) 
commdists <- st_as_sf(commdists) %>% 
  mutate(NAME = ifelse(DISTRICT == "1", "1 North",
                ifelse(DISTRICT == "2", "2 South", "3 Central")))
the_commdists <- unique(st_set_geometry(commdists, NULL)$NAME)
loop_commdists <- st_as_sf(commdists) %>% mutate(tomatch = NAME)

uga_no_city <- st_as_sf(uga) %>%
  filter(substring(GMA_JURISD, 1, 4) != "City")
the_ugas <- unique(st_set_geometry(uga_no_city, NULL)$GMA_JURISD)
loop_ugas <- st_as_sf(uga_no_city) %>% mutate(tomatch = GMA_JURISD)

the_lamirds <- unique(st_set_geometry(lamird, NULL)$PLAN_AREA)
loop_lamirds <- st_as_sf(lamird) %>% mutate(tomatch = PLAN_AREA)

pub_schools <- schools %>%
  filter(TYPE == "PUBLIC")
pub_schools2 <- getWalkingPolygonInCounty(pub_schools, "NAME")
the_schools <- unique(st_set_geometry(pub_schools2, NULL)$id)
loop_schools <- st_as_sf(pub_schools2) %>% mutate(tomatch = id)

libraries <- community %>%
  filter(TYPE == "PUBLIC LIBRARY")
libraries2 <- getWalkingPolygonInCounty(libraries, "NAME")
the_libraries <- unique(st_set_geometry(libraries2, NULL)$id)
loop_libraries <- st_as_sf(libraries2) %>% mutate(tomatch = id)

transit_cp <- tribble(
  ~NAME, ~ADDRESS,
  "Bremerton Transportation Center (Ferry Terminal)","10 Washington Ave, Bremerton 98337",
  "Bainbridge Island Ferry Terminal","270 Olympic Drive SE, Bainbridge 98110",
  "Wheaton Way Transit Center","3915 Wheaton Way, Bremerton 98310",
  "West Bremerton Transit Center","540 Bruenn Ave, Bremerton 98312",
  "Silverdale Transit Center","Greaves Way & Kitsap Mall Blvd, Silverdale 98383",
  "North Viking Transit Center","21992 Viking Ave NW, Poulsbo 98370",
  "Kingston Ferry Terminal","11264 State Route 104, Kingston 98346",
  "Port Orchard Ferry Dock","73 Sidney Ave, Port Orchard 98366",
  "Southworth Ferry Terminal","11564 SE State Hwy. 160, Southworth 98386",
  "Annapolis Ferry Dock","1076 Bay Street, Port Orchard 98366",
  "Hwy. 305 & Suquamish Way","16003 WA-305, Poulsbo, Washington 98370",
  "McWilliams Park & Ride","1601 NE McWilliams Rd, Bremerton, Washington 98311",
  "Suquamish Park & Ride","18829 Division Ave NE, Suquamish, Washington 98392",
  "Port Orchard Wal-Mart","3497 Bethel Rd SE, Port Orchard 98366",
  "George's Corner Park & Ride","27618 Hansville Rd NE, Kingston 98346",
  "Gateway Fellowship Park & Ride","18901 8th Ave NE, Poulsbo 98370",
  "Miller Bay & Indianola","23404 Miller Bay Rd NE Poulsbo, Washington 98370",
  "St. Gabriel's Church","1150 Mitchell Ave SE, Port Orchard 98366"
)
# geocode the addresses into geocode = c(lon, lat)
transit_cp <- transit_cp %>% #tmaptools::geocode_OSM(transit_cp$ADDRESS)
  rowwise() %>%
  mutate(geocode = list(set_names(mapboxapi::mb_geocode(ADDRESS),
                        c("lon","lat")))) %>%
  unnest_wider(geocode)

# convert the lat lon into a point geometry
transits <- st_as_sf(transit_cp, 
                      coords = c(x = "lon", y = "lat"), 
                      crs = 4326) 
transits2 <- getWalkingPolygonInCounty(transits, "NAME")
the_transits <- unique(st_set_geometry(transits2, NULL)$id)
loop_transits <- st_as_sf(transits2) %>% mutate(tomatch = id)

# combine data into tibble for this report
#These regions and pedestrian generators are:
kds <- "Kitsap County GIS"
kt <- "Kitsap Transit Website"
wd <- "Walking distance polygons from Mapbox"
breakdowns <- tribble(
  ~Name,~pointdf,~polydf,~idcolpoly,~items,~data_source,
  "Commissioner Districts",NA,"commdists","NAME","the_commdists",kds,
  "UGAs",NA,"uga_no_city","GMA_JURISD","the_ugas",kds,
  "LAMIRDs",NA,"lamird","PLAN_AREA","the_lamirds",kds,
  "Public Schools","pub_schools","pub_schools2","id","the_schools",paste(kds,wd),
  "Public Libraries","libraries","libraries2","id","the_libraries",paste(kds,wd),
  "Transit Centers","transits","transits2","id","the_transits",paste(kt,wd),
  )

@@

colnames(groupcl) <- c("Road Classification", 
                       "Count of Roads", "Count of Segments", 
                       "Length in Miles") 
knitr::kable(groupcl, 
             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

knitr::kable(df, 
             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

knitr::kable(df, 
             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

```{r combine-rd-sw-sh}
# combine the sidewalk and shoulder and roads data
x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 600)
x$RD_LOG_ID <- 43809
comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
                                     SEGMENT_ID_R = SEGMENT_ID,
                                     FULL_NAME_R = FULL_NAME, 
                                     LENGTH_R = LENGTH,
                                     road_class, RUCODE), 
                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
  mutate(distance = as.integer(st_length(.)))

@@

knitr::kable(example_tbl) %>%
kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
              full_width = F, 
              position = "float_left")
```

@@

df <- score_interpretation %>%
  mutate(`Score At Least` = ifelse(score_at_least == 0.001, "> 0", 
                                   nice_comma(score_at_least,0))) %>%
  select(`Score At Least`, Interpretation = interpretation)

knitr::kable(df,
             caption = "Interpretation of Scores for Segments and for Areas")  %>%
   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                full_width = F)
```

@@

comb_score2 <- st_set_geometry(ncdf, NULL) %>% 
  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>%
  left_join(comb_score) %>%
  mutate(the_score = ifelse(is.na(score), 0, score),
         the_length = ifelse(is.na(distance), LENGTH, distance),
         w_score = the_score*the_length)

@@

```{r tbl_combscorestandard}
knitr::kable(comb_score_by_setting_class %>%
               rename(Setting = setting,
                      Class = road_class,
                      Score = score,
                      Interpretation = interpretation) %>%
               mutate(Score = ifelse(Score == 0, "0", nice_comma(Score, 1))),
             caption = "Weighted Scores by Standard Group")  %>%
   footnote(general = "Unknown setting is scored as if it is Urban."
                      ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r map_score}

comb_map <- st_as_sf(ncdf) %>% 
  select(RD_LOG_ID_R = RD_LOG_ID, SEGMENT_ID_R = SEGMENT_ID,
         FULL_NAME_R = FULL_NAME, road_class, RUCODE, LENGTH) %>%
  mutate(setting = ifelse(RUCODE == 1, "Rural", 
                          ifelse(RUCODE == 2, "Urban", "Unknown"))) %>% 
  left_join(comb_score) %>%
  rowwise() %>%
  mutate(interpretation = getScoreInterpretation(score))

score_col <- scale_color_manual(values = c("Good"="#4d9221",
                                           "Adequate"="#a1d76a",
                                           "Minimum" = "#e6f5d0",
                                           "Poor" = "#fde0ef",
                                           "Very Poor" = "#e9a3c9",
                                           "Nothing" = "#c51b7d"),
                                na.value = "#808080")
comb_map$interpretation <- factor(comb_map$interpretation, 
                                  levels =  score_interpretation$interpretation)
kitsap + 
  geom_sf(data = cities["NAME"], 
          aes(fill="NAME"),
          fill="#808080",
          alpha = 0.5,
          show.legend = FALSE) +
  geom_sf(data = ncdf, aes(color = 1), color = "#808080", show.legend = FALSE) + 
  geom_sf(data = comb_map, aes(color = interpretation), show.legend = "line") + 
  score_col +
  theme_void() +
  labs(title = "Calculated Pedestrian Facility Score",
       caption = "Data source: Kitsap GIS, scores calculated",
       color = "Score Interpretation")

@@

Some of the segments with score of 0 are because they have shoulders when they should have sidewalks. If you are familiar with Central Kitsap, you might recognize this kind of situation on Provost/Old Frontier - where there are wide shoulders, but the road is in the urban growth area and identified as urban; therefore, design standards suggest sidewalks are appropriate. Here's a few example segments.

@@

Some road segments are missing entirely from the centerline road data - for example, the portion of Ridgetop that goes from the end of Myhre and curves back to Silverdale Way. These missing road segments on Ridgetop are mostly complete streets with wide sidewalks. Here, showing a map with all roads with the name *RIDGETOP*, you can only see the part going from Silverdale Way to Myhre.^[For completeness, we also checked the road log id, and it has the same data.] These segments, and any others like them, missing from the data decrease the overall pedestrian facilities score.

@@

We can also look at regions and pedestrian generators within the County. For these, we identify the roads and any sidewalks or shoulders.^[In theory, the miles of recorded sidewalk or shoulder and matched miles could be double the centerline miles if every mile had a sidewalk or shoulder.]

@@

These regions and pedestrian generators are:

```{r, results = 'asis'}
cat(paste0("- ", breakdowns$Name, collapse = "\n"))
```

For each smaller region or pedestrian generator type, we show a map of Kitsap County (sans incorporated cities) indicating locations of the regions of that type.

Then, for each individual region, we show the pedestrian facilities on the left side and the right side of the road and calculate scores.

```{r}
# this starts the tibble that will hold ALL the scores
score_breakdowns <- tibble(Area = "Kitsap County", Score = comb_total_score)
```

```{r bdlooping, include=FALSE, echo=FALSE}

expanded_region <- lapply(
    seq_along(1:nrow(breakdowns)),
    function(irow) {
      knitr::knit_expand(file = "child_region.Rmd",
                                   i = irow )
    })

parsed_region <- knitr::knit_child(text = unlist(expanded_region))
```

`r parsed_region`

# Conclusion

@@

Smaller breakdowns of the county fare differently. Here, we show a table for each region type. Any region with no segments getting more than a 0, "Nothing" show up with red text. The best region of that type will have bold text, and any regions with a score of **Adequate** will have green text.

@@

getScoreBreakdownForArea <- function(the_area) {
  # get name
  the_area_name <- breakdowns$Name[breakdowns$items == the_area]
  the_area_df <- score_breakdowns %>% 
                 filter(Area %in% eval(as.symbol(the_area))) %>%
                 mutate(s = Score,
                        Score = ifelse(Score == 0,
                                       nice_comma(Score,0),
                                       nice_comma(Score,2))) %>%
                 arrange(Area)
  knitr::kable( select(the_area_df, -s),
                caption = paste("Scores for", the_area_name )
             ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  row_spec(which(the_area_df$s == 0), color = "red") %>%
  row_spec(which(the_area_df$s == max(the_area_df$s)), bold = T) %>%
  row_spec(which(the_area_df$Interpretation == "Adequate"), bold = T, color = "green")
}

@@

```{r, echo=FALSE, results='asis'}
score_breakdowns <- read_csv(paste0(this_dir,"score_breakdowns.csv"))
res <- lapply(breakdowns$items, function(x) {
  knitr::knit_child(text = c(
    '',
    '```{r}',
    'getScoreBreakdownForArea(x)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')

@@

 * missing part of Ridgetop Blvd in Silverdale where sidewalks are known to exist
 * missing data on sidewalks that are known to exist
 * unmatched road log identifiers between road and sidewalk and shoulder data

@@

In addition, many road segments have shoulders when their standards suggest they should have sidewalks. Is the presence of a shoulder, though not adequate, really the same as **nothing?** - that is how the scoring currently calculates.

One cautionary note - while this method evaluates pedestrian facilities, it is entirely focused on the presence of facilities of adequate width and separation from cars. It is not taking into consideration the quality of those facilities. Disrepair (cracks, uneven surfaces) and encroaching vegetation can greatly reduce the usability of pedestrian facilities. Quality of facilities is beyond the scope of this evaluation.

@@

# References
